---
title: "Ottenby_WeightFat"
author: "Maria Laura Mahecha Escobar"
format: pdf
editor: visual
---

## Data Set up

Libraries we need

```{r libraries}
library(stringr)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(performance)
library(dplyr)
library(lme4)
library(ggpmisc)
```

First set directory

```{r Directory}
#| include: false 
setwd("E:/Users/mlmah/OneDrive/Documentos/MLME/Maestria/Animal_Ecology_2022_2024/Tesis/Database/Autumn/Stensoffa/Stensoffa_activity_copia/2023_Robin_MariaLaura")
```

Now import the dataset

```{r Mass_df}
#| output: false
#| include: false
Mass_data<-read.table("E:/Users/mlmah/OneDrive/Documentos/MLME/Maestria/Animal_Ecology_2022_2024/Tesis/Database/Weight_Spring.txt", dec=",", h=T)
```

We can plot the mass of the individuals across days per spp

```{r plot_basic_mass}
ggplot(Mass_data, aes(x=ExperimentalDay, y=Mass, colour=Species))+
  geom_point()+
  geom_smooth(method='lm')
```

We can also see the body condition (mass/wing length)

```{r plot_basi_cond}
ggplot(Mass_data, aes(x=ExperimentalDay, y=BodyCond, colour=Species))+
  geom_point()+
  geom_smooth(method='lm')
```

And to check that body condition can be used instead of mass

```{r BodyCond_Mass_plot}
ggplot(Mass_data, aes(x=Mass, y=BodyCond, colour=Species))+
  geom_point()+
  geom_smooth(method="lm")
```

And just to check that mass and fat are related. First we calculate the change in mass and the change in fat

```{r Diff_massFat}
Mass_data$Mass_diff <- c(NA, diff(Mass_data$Mass))
Mass_data$Fat_diff <- c(NA, diff(Mass_data$Fat))
```

Now we can plot the difference

```{r plot_diff_MassFat}
ggplot(Mass_data, aes(x=Fat_diff, y=Mass_diff, colour=Species))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)

ggplot(Mass_data, aes(x=Fat, y=Mass, colour=Species))+
     geom_point()+
     geom_hline(yintercept = 0)+
     geom_vline(xintercept = 0)+
     stat_poly_line() +
     stat_poly_eq(use_label(c("adj.R2","p", "n"))) +
     facet_wrap(vars(Species))
```

### Lmer

#### BodyCond\~Treat

```{r mass_mod1}
Mass_mod1<-lmer(BodyCond ~ Treatment + (1|Ring) + (1|Species), data= Mass_data,
                na.action = na.exclude)

check_model(Mass_mod1)

summary(Mass_mod1)

mean_Mass_mod1<-emmeans(Mass_mod1, ~Treatment)
mean_Mass_mod1

pairs(mean_Mass_mod1)

gr_Mass_mod1 <- ref_grid(Mass_mod1, cov.keep=c("Treatment"))
gr_emmeans_Mass_mod1<-as.data.frame(emmeans(gr_Mass_mod1, 
                                            specs = c("Treatment"), 
                                            level=0.95))

mean_Mass_mod1<- as.data.frame(mean_Mass_mod1)

raw_Mass_mod1 <- Mass_data[c("Treatment", "BodyCond")]  

plot_data_Mass_mod1 <- merge(gr_emmeans_Mass_mod1, raw_Mass_mod1,
                        by = c("Treatment")) 

ggplot(mean_Mass_mod1, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Mass_mod1, aes(x = Treatment, y = BodyCond),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  #facet_wrap(vars(Species))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Mass_mod1)

null_mass_mod1<- lmer(BodyCond ~ 1 + (1|Ring) + (1|Species), data= Mass_data,
                na.action = na.exclude)
anova(Mass_mod1, null_mass_mod1, test="LTR")
```

#### Fat \~ treat

```{r fat_mod1}
Fat_mod1<-lmer(Fat ~ Treatment + (1|Ring) + (1|Species), data= Mass_data,
                na.action = na.exclude)

check_model(Fat_mod1)

summary(Fat_mod1)

mean_Fat_mod1<-emmeans(Fat_mod1, ~Treatment)
mean_Fat_mod1

pairs(mean_Fat_mod1)

gr_Fat_mod1 <- ref_grid(Fat_mod1, cov.keep=c("Treatment"))
gr_emmeans_Fat_mod1<-as.data.frame(emmeans(gr_Fat_mod1, 
                                            specs = c("Treatment"), 
                                            level=0.95))

mean_Fat_mod1<- as.data.frame(mean_Fat_mod1)

raw_Fat_mod1 <- Mass_data[c("Treatment", "Fat")]  

plot_data_Fat_mod1 <- merge(gr_emmeans_Fat_mod1, raw_Fat_mod1,
                        by = c("Treatment")) 

ggplot(mean_Fat_mod1, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Fat_mod1, aes(x = Treatment, y = Fat),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  #facet_wrap(vars(Species))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Fat_mod1)

null_Fat_mod1<- lmer(Fat ~ 1 + (1|Ring) + (1|Species), data= Mass_data,
                na.action = na.exclude)
anova(Fat_mod1, null_Fat_mod1, test="LTR")
```

#### BodyCond\~Treatment\*BreedingDistribution

```{r mass_mod2}
Mass_mod2<-lmer(BodyCond ~ Treatment*BreedingDistribution + 
                  (1|Ring) + (1|Species), data= Mass_data,
                na.action = na.exclude)

check_model(Mass_mod2)

summary(Mass_mod2)

mean_Mass_mod2<-emmeans(Mass_mod2, ~Treatment*BreedingDistribution)
mean_Mass_mod2

pairs(mean_Mass_mod2)

gr_Mass_mod2 <- ref_grid(Mass_mod2, cov.keep=
                           c("Treatment", "BreedingDistribution"))
gr_emmeans_Mass_mod2<-as.data.frame(emmeans(gr_Mass_mod2, 
                                            specs = c("Treatment",
                                                      "BreedingDistribution"), 
                                            level=0.95))

mean_Mass_mod2<- as.data.frame(mean_Mass_mod2)

raw_Mass_mod2 <- Mass_data[c("Treatment", "BodyCond","BreedingDistribution")]  

plot_data_Mass_mod2 <- merge(gr_emmeans_Mass_mod2, raw_Mass_mod2,
                        by = c("Treatment","BreedingDistribution")) 

ggplot(mean_Mass_mod2, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Mass_mod2, aes(x = Treatment, y = BodyCond),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  facet_wrap(vars(BreedingDistribution))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Mass_mod2)

null_mass_mod1<- lmer(BodyCond ~ 1 + (1|Ring) + (1|Species), data= Mass_data,
                na.action = na.exclude)
anova(Mass_mod1, Mass_mod2, null_mass_mod1, test="LTR")
```

#### Fat\~Treatment\*BreedingDistribution

```{r fat_mod2}
Fat_mod2<-lmer(Fat ~ Treatment*BreedingDistribution + 
                  (1|Ring) + (1|Species), data= Mass_data)

check_model(Fat_mod2)

summary(Fat_mod2)

mean_Fat_mod2<-emmeans(Fat_mod2, ~ Treatment*BreedingDistribution)
mean_Fat_mod2

pairs(mean_Fat_mod2)

gr_Fat_mod2 <- ref_grid(Fat_mod2, cov.keep=
                           c("Treatment", "BreedingDistribution"))
gr_emmeans_Fat_mod2<-as.data.frame(emmeans(gr_Fat_mod2, 
                                            specs = c("Treatment",
                                                      "BreedingDistribution"), 
                                            level=0.95))

mean_Fat_mod2<- as.data.frame(mean_Fat_mod2)

raw_Fat_mod2 <- Mass_data[c("Treatment", "Fat", "BreedingDistribution")]  

plot_data_Fat_mod2 <- merge(gr_emmeans_Fat_mod2, raw_Fat_mod2,
                        by = c("Treatment","BreedingDistribution")) 

ggplot(mean_Fat_mod2, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Fat_mod2, aes(x = Treatment, y = Fat),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  facet_wrap(vars(BreedingDistribution))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Fat_mod2)

null_Fat_mod1<- lmer(Fat ~ 1 + (1|Ring) + (1|Species), data= Mass_data,
                na.action = na.exclude)
anova(Fat_mod1, Fat_mod2, null_Fat_mod1, test="LTR")
```

## Robins

Set the dataset for Robins

```{r Mass_Robins}
Mass_data_Robin<- subset(Mass_data, Species=="Robin")
```

Basic plot

```{r plot_basic_mass_R}
ggplot(Mass_data_Robin, aes(x=ExperimentalDay, y=Mass))+
  geom_point()+
  geom_smooth(method='lm')
```

Plotting condition

```{r plot_basi_cond_R}
ggplot(Mass_data_Robin, aes(x=ExperimentalDay, y=BodyCond))+
  geom_point()+
  geom_smooth(method='lm')
```

Check the fat difference against mass change

```{r Diff_massFat_R}
Mass_data_Robin$Mass_diff <- c(NA, diff(Mass_data_Robin$Mass))
Mass_data_Robin$Fat_diff <- c(NA, diff(Mass_data_Robin$Fat))
```

Now plot the difference

```{r plot_diff_MassFat_R}
ggplot(Mass_data_Robin, aes(x=Fat_diff, y=Mass_diff))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)
```

### Lmer

#### BodyCond\~Treatment

```{r}
Mass_mod1_R<-lmer(BodyCond ~ Treatment + (1|Ring), data= Mass_data_Robin,
                na.action = na.exclude)

check_model(Mass_mod1_R)

summary(Mass_mod1_R)

mean_Mass_mod1_R<-emmeans(Mass_mod1_R, ~Treatment)
mean_Mass_mod1_R

pairs(mean_Mass_mod1_R)

gr_Mass_mod1_R <- ref_grid(Mass_mod1_R, cov.keep=c("Treatment"))
gr_emmeans_Mass_mod1_R<-as.data.frame(emmeans(gr_Mass_mod1_R, 
                                            specs = c("Treatment"), 
                                            level=0.95))

mean_Mass_mod1_R<- as.data.frame(mean_Mass_mod1_R)

raw_Mass_mod1_R <- Mass_data_Robin[c("Treatment", "BodyCond")]  

plot_data_Mass_mod1_R <- merge(gr_emmeans_Mass_mod1_R, raw_Mass_mod1_R,
                        by = c("Treatment")) 

ggplot(mean_Mass_mod1_R, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Mass_mod1_R, aes(x = Treatment, y = BodyCond),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  #facet_wrap(vars(Species))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Mass_mod1_R)

null_mass_mod1_R<- lmer(BodyCond ~ 1 + (1|Ring), 
                        data= Mass_data_Robin,
                na.action = na.exclude)
anova(Mass_mod1_R, null_mass_mod1_R, test="LTR")
```

## Willys

Set the dataset for Willys

```{r Mass_Willys}
Mass_data_Willys <- subset(Mass_data, Species=="Willow_Warbler")
```

Basi plot

```{r plot_basic_mass_W}
ggplot(Mass_data_Willys, aes(x=ExperimentalDay, y=Mass))+
  geom_point()+
  geom_smooth(method='lm')
```

ploting condition

```{r plot_basi_cond_W}
ggplot(Mass_data_Willys, aes(x=ExperimentalDay, y=BodyCond))+
  geom_point()+
  geom_smooth(method='lm')
```

Check the fat difference against mass change

```{r Diff_massFat_W}
Mass_data_Willys$Mass_diff <- c(NA, diff(Mass_data_Willys$Mass))
Mass_data_Willys$Fat_diff <- c(NA, diff(Mass_data_Willys$Fat))
```

Now plot the difference

```{r plot_diff_MassFat_W}
ggplot(Mass_data_Willys, aes(x=Fat_diff, y=Mass_diff))+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)
```

### Lmer

#### BodyCond\~Treatment

```{r mass_mod1_W}
Mass_mod1_W<-lmer(BodyCond ~ Treatment + (1|Ring), data= Mass_data_Willys,
                na.action = na.exclude)

check_model(Mass_mod1_W)

summary(Mass_mod1_W)

mean_Mass_mod1_W<-emmeans(Mass_mod1_W, ~Treatment)
mean_Mass_mod1_W

pairs(mean_Mass_mod1_W)

gr_Mass_mod1_W <- ref_grid(Mass_mod1_W, cov.keep=c("Treatment"))
gr_emmeans_Mass_mod1_W<-as.data.frame(emmeans(gr_Mass_mod1_W, 
                                            specs = c("Treatment"), 
                                            level=0.95))

mean_Mass_mod1_W<- as.data.frame(mean_Mass_mod1_W)

raw_Mass_mod1_W <- Mass_data_Willys[c("Treatment", "BodyCond")]  

plot_data_Mass_mod1_W <- merge(gr_emmeans_Mass_mod1_W, raw_Mass_mod1_W,
                        by = c("Treatment")) 

ggplot(mean_Mass_mod1_W, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Mass_mod1_W, aes(x = Treatment, y = BodyCond),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  #facet_wrap(vars(Species))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Mass_mod1_W)

null_mass_mod1_W<- lmer(BodyCond ~ 1 + (1|Ring), 
                        data= Mass_data_Willys,
                na.action = na.exclude)
anova(Mass_mod1_W, null_mass_mod1_W, test="LTR")
```
