---
title: "Emlen_Ottenby"
author: "Maria Laura Mahecha Escobar"
format: pdf
editor: visual
---

## 

```{r}
#| output: FALSE 
setwd("E:/Users/mlmah/OneDrive/Documentos/MLME/Maestria/Animal_Ecology_2022_2024/Tesis/Database")
```

For running this script you need the following packages installed in your workspace:

```{r libraries, output=FALSE, include=TRUE, warning=FALSE}
library(usethis) #To use github
library(knitr) #to use markdown
library(circular) #to perform circular statistics
library(ggplot2) #to do beautiful plots
library(rmarkdown) #to use markdown
library(quarto) #to use markdown
library(lmerTest) #to do random effect tests
library(cowplot)
library(lme4) #to do random effect tests
library(dplyr) #managing tables and organizing data
library(janitor)
library(tvthemes)
library(phytools) #managing phylogenetic approaches
library(bpnreg) #circular regression models
library(CircStats) #some useful circular functions
library(RSQLite) #to connect database reades
library(data.table) #to handle tables
library(remotes)
library(tidyr)
library(readr)
library(CircStats)
library(performance)
library(tectonicr)
```

## Emlen funnel analysis

Lets read the table with the Emlen data

```{r database}
#| output: false
#| include: false
Emlen_IndData<-read.table("E:/Users/mlmah/OneDrive/Documentos/MLME/Maestria/Animal_Ecology_2022_2024/Tesis/Database/EmlenData_Spring2023.txt", h=T)
```

Let's change the names of the columns for the middle of the angles the sector has

```{r newSectorNames}
setnames(Emlen_IndData, old=c("S1","S2","S3","S4","S5","S6","S7","S8","S9",
                              "S10","S11","S12","S13","S14","S15","S16","S17",
                              "S18","S19","S20","S21","S22","S23","S24"), 
         new = c("7.5","22.5","37.5","52.5","67.5","82.5","97.5","112.5",
                 "127.5","142.5","157.5","172.5","187.5","202.5","217.5",
                 "232.5","247.5","262.5","277.5","292.5","307.5","322.5",
                 "337.5","352.5"), skip_absent=TRUE)

```

To add an identification to every Emlen experiment, because there are more than one experiment per ring

```{r DirectionTable}
Emlen_IndData<-tibble::rowid_to_column(Emlen_IndData, "Identificator")
```

Let's define two functions that are going to be helpful handling circular data because there's functions that need the data input in degrees and others in radians. With this functions I can change it easily

```{r DegRad_Functions}
deg2rad <- function(deg) {(deg * pi) / (180)} #converts degrees to radians.
rad2deg <- function(rad) {(rad * 180) / (pi)} #converts radians to degrees.
```

Lets organize the dates in the table

```{r Weeks}
Emlen_IndData$Date<-lubridate::dmy(Emlen_IndData$Date)
Emlen_IndData <- Emlen_IndData %>%
  mutate(Week = case_when(
    between(Date, as.Date("2023-04-20"), as.Date("2023-04-26")) ~ 1,
    between(Date, as.Date("2023-04-28"), as.Date("2023-05-04")) ~ 2,
    between(Date, as.Date("2023-05-08"), as.Date("2023-05-14")) ~ 3,
    between(Date, as.Date("2023-05-16"), as.Date("2023-05-22")) ~ 4,
    between(Date, as.Date("2023-05-25"), as.Date("2023-05-31")) ~ 5,
    TRUE ~ 6 # Defining values for each week with date intervals
  ))
```

This code makes the reppetition of the angles to a long format to being able to analyze the data after

```{r data_long}
datalong_EmlenInd<-pivot_longer( #makes it a frequency to rep after
  Emlen_IndData,
  cols=c("7.5","22.5","37.5","52.5","67.5","82.5","97.5","112.5","127.5",
         "142.5","157.5","172.5","187.5","202.5","217.5","232.5","247.5",
         "262.5","277.5","292.5","307.5","322.5","337.5","352.5"),
  names_to="sector",
  values_to="count")
```

Lets save the all the numbers in one variable to make a new table with the

```{r uniqueIdent}
uniqueValues_ID<-unique(datalong_EmlenInd$Identificator)
```

This is a really important piece of code because I'm going to store the statistics results for each individual experiment

```{r Results_df}
#| warning: false
results_df <- data.frame(Identificator = uniqueValues_ID,   #create an empty dataframe to store results
                         test_statistic = numeric(length(uniqueValues_ID)),
                         p_value = numeric(length(uniqueValues_ID)),
                         Mean = numeric(length(uniqueValues_ID)),
                         Lower_CI = numeric(length(uniqueValues_ID)),
                         Upper_CI = numeric(length(uniqueValues_ID)))

# Loop through unique identifiers
for (i in uniqueValues_ID) {
  subset_data <- subset(datalong_EmlenInd, Identificator == i) #subsets data for each identificator/each entry in the database
  IndData <- as.data.frame(as.numeric(unlist(subset_data)))
  IndData <- na.omit(IndData)
  longdata <- as.numeric(rep(subset_data$sector, times = subset_data$count)) #repeats the degrees the times we counted already
  conf_interval <- confidence_interval(longdata, conf.level = 0.95, axial = FALSE)
  longdata <- deg2rad(longdata)  # Convert to radians
  rayleigh_test <- rayleigh.test(longdata)#performs rayleigh test for each entry in the database
  test_statistic <- rayleigh_test$statistic
  p_value <- rayleigh_test$p.value
  MeanDir<- mean.circular(longdata)
  
  
  # Convert the mean direction from radians to degrees and correct negative values
  MeanDir_deg <- rad2deg(MeanDir)  # Convert to degrees
  if (MeanDir_deg < 0) {
    MeanDir_deg <- 360 + MeanDir_deg  # Wrap around negative values
  }
  
  # Assign results to the correct rows
  row_index <- which(results_df$Identificator == i)
  results_df$test_statistic[row_index] <- test_statistic
  results_df$p_value[row_index] <- p_value
  results_df$Mean[row_index]<- MeanDir_deg
  results_df$Lower_CI[row_index]<- conf_interval$conf.interval[1]
  results_df$Upper_CI[row_index]<- conf_interval$conf.interval[2]
}

```

We have the results per identificator. Now we merge both tables to have all the data associated to each one of the rings

```{r mergedEmlen}
mergedEmlen<-merge(results_df, Emlen_IndData, by = "Identificator") #merge both dataframes by identificator
```

To check unique values and how many data points a ring has to perform a paired test later

```{r uniqueValues_Ring}
uniqueValues_Ring<-unique(mergedEmlen$Ring)
length(uniqueValues_Ring)
```

to see how many times a ring occours in the dataframe (this is because we need to have the same size if we want to do a paired test)

```{r occurrences}
occurrences <- mergedEmlen %>% count(Ring) 
print(occurrences)
```

Now we can choose the columns to make the tests

```{r EmlenData_ForTest}
EmlenData_ForTest<- dplyr::select(mergedEmlen, c("Identificator", "test_statistic","p_value", "Mean", "Lower_CI", "Upper_CI", "Species", "Treatment","Ring","BreedingDistribution", "Week"))
```
