---
title: "Stensoffa_WeightFat"
author: "Maria Laura Mahecha Escobar"
format: pdf
editor: visual
---

## 

## Data Set up

Libraries we need

```{r libraries}
library(stringr) 
library(ggplot2) 
library(ggpubr) 
library(usethis) #To use github
library(knitr) #to use markdown
library(rmarkdown) #to use markdown
library(quarto) #to use markdown
library(lmerTest) #to do random effect tests
library(lme4) #to do random effect tests
library(dplyr) #managing tables and organizing data
library(janitor)
library(tvthemes)
library(phytools) #managing phylogenetic approaches
library(bpnreg) #circular regression models
library(CircStats) #some useful circular functions
library(RSQLite) #to connect database reades
library(data.table) #to handle tables
library(tidyr)
library(readr)
library(performance)
library(tectonicr)
library(wesanderson)
library(fitdistrplus)
library(visreg)
library(emmeans)
```

First set directory

```{r Directory}
#| include: false  
setwd("E:/Users/mlmah/OneDrive/Documentos/MLME/Maestria/Animal_Ecology_2022_2024/Tesis/Database/Autumn/Stensoffa/Stensoffa_activity_copia/2023_Robin_MariaLaura")
```

Now import the dataset

```{r Mass_df}
#| output: false 
#| include: false 
stf_Mass_data<-read.table("E:/Users/mlmah/OneDrive/Documentos/MLME/Maestria/Animal_Ecology_2022_2024/Tesis/Database/Weight_Autumn.txt", dec=",", h=T)
```

We can plot the mass of the individuals across days per spp

```{r plot_basic_mass}
ggplot(stf_Mass_data, aes(x=ExperimentalDay, y=Mass, colour=Treatment))+  
  geom_point()+   geom_smooth(method='lm')
```

We can also see the body condition (mass/wing length)

```{r plot_basi_cond}
ggplot(stf_Mass_data, aes(x=ExperimentalDay, y=BodyCond, colour=Treatment))+   
  geom_point()+   
  geom_smooth(method='lm')
```

And to check that body condition can be used instead of mass

```{r BodyCond_Mass_plot}
ggplot(stf_Mass_data, aes(x=Mass, y=BodyCond, colour=Treatment))+  
  geom_point()
```

Lets check the normality of bodycond

```{r}
hist(stf_Mass_data$BodyCond)
shapiro.test(stf_Mass_data$BodyCond)
```

### Lmer

#### BodyCond\~Treat

```{r mass_mod1}
#MIRAR EL MATERIAL SUPLEMENTARIO DE LOS DUNNOKS PARA VER EL RANDOM INTERCEPT OR SLOPE- CHECK!
Mass_mod1_stf<-lmer(BodyCond ~ Treatment + (1|Ring), 
                    data= stf_Mass_data,
                    na.action = na.exclude)

check_model(Mass_mod1_stf)

summary(Mass_mod1_stf)

mean_Mass_mod1_stf<-emmeans(Mass_mod1_stf, ~Treatment)
mean_Mass_mod1_stf

pairs(mean_Mass_mod1_stf)

gr_Mass_mod1_stf <- ref_grid(Mass_mod1_stf, cov.keep=c("Treatment"))
gr_emmeans_Mass_mod1_stf<-as.data.frame(emmeans(gr_Mass_mod1_stf, 
                                            specs = c("Treatment"), 
                                            level=0.95))

mean_Mass_mod1_stf<- as.data.frame(mean_Mass_mod1_stf)

raw_Mass_mod1_stf <- stf_Mass_data[c("Treatment", "BodyCond")]  

plot_data_Mass_mod1_stf <- merge(gr_emmeans_Mass_mod1_stf, raw_Mass_mod1_stf,
                        by = c("Treatment")) 

ggplot(mean_Mass_mod1_stf, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Mass_mod1_stf, aes(x = Treatment, y = BodyCond),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  #facet_wrap(vars(Species))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Mass_mod1_stf)

null_mass_mod1_stf<- lmer(BodyCond ~ 1 + (1|Ring), 
                        data= stf_Mass_data,
                na.action = na.exclude)
anova(Mass_mod1_stf, null_mass_mod1_stf, test="LTR")
```

#### BodyCond\~Treatment\*ExperimentalDay

```{r}
Mass_mod2_stf<-lmer(BodyCond ~ Treatment*ExperimentalDay + (1|Ring),
                    data= stf_Mass_data,
                    na.action = na.exclude)

check_model(Mass_mod2_stf)

summary(Mass_mod2_stf)

mean_Mass_mod2_stf<-emmeans(Mass_mod2_stf, ~Treatment*ExperimentalDay)
mean_Mass_mod2_stf

pairs(mean_Mass_mod2_stf)

gr_Mass_mod2_stf <- ref_grid(Mass_mod2_stf, cov.keep=
                               c("Treatment", "ExperimentalDay"))
gr_emmeans_Mass_mod2_stf<-as.data.frame(emmeans(gr_Mass_mod2_stf, 
                                            specs = 
                                              c("Treatment", "ExperimentalDay"), 
                                            level=0.95))

mean_Mass_mod2_stf<- as.data.frame(mean_Mass_mod2_stf)

raw_Mass_mod2_stf <- stf_Mass_data[c("Treatment", "BodyCond", "ExperimentalDay")] 

stf_Mass_pval_mod2<-emmeans(gr_Mass_mod2_stf, specs = 
                              c("Treatment", "ExperimentalDay"),
                       at = list(Days = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14)),
                       level=0.95)
stf_Mass_pval_mod2_df<-as.data.frame(pairs(stf_Mass_pval_mod2, adjust=NULL))

plot_data_Mass_mod2_stf <- merge(gr_emmeans_Mass_mod2_stf, raw_Mass_mod2_stf,
                        by = c("Treatment", "ExperimentalDay")) 

ggplot(gr_emmeans_Mass_mod2_stf, 
       aes(x = ExperimentalDay, y = emmean, colour=Treatment)) +
  geom_jitter(data = plot_data_Mass_mod2_stf, 
              aes(x = ExperimentalDay, y = BodyCond),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5)+
  geom_line()+
  facet_wrap(vars(Treatment))+
  labs(x = "ExperimentalDay", y = "Estimated Marginal Mean") +
  theme_bw()

model_performance(Mass_mod2_stf)

anova(Mass_mod2_stf, Mass_mod1_stf, null_mass_mod1_stf, test="LTR")
```

#### FoodEaten\~Treat

```{r mod3_stf}
Mass_mod3_stf<-lmer(FoodEaten ~ Treatment + (1|Ring), data= stf_Mass_data,
                na.action = na.exclude)

check_model(Mass_mod3_stf)

summary(Mass_mod3_stf)

mean_Mass_mod3_stf<-emmeans(Mass_mod3_stf, ~Treatment)
mean_Mass_mod3_stf

pairs(mean_Mass_mod3_stf)

gr_Mass_mod3_stf <- ref_grid(Mass_mod3_stf, cov.keep=c("Treatment"))
gr_emmeans_Mass_mod3_stf<-as.data.frame(emmeans(gr_Mass_mod3_stf, 
                                            specs = c("Treatment"), 
                                            level=0.95))

mean_Mass_mod3_stf<- as.data.frame(mean_Mass_mod3_stf)

raw_Mass_mod3_stf <- stf_Mass_data[c("Treatment", "FoodEaten")]  

plot_data_Mass_mod3_stf <- merge(gr_emmeans_Mass_mod3_stf, raw_Mass_mod3_stf,
                        by = c("Treatment")) 

ggplot(mean_Mass_mod3_stf, aes(x = Treatment, y = emmean)) +
  geom_jitter(data = plot_data_Mass_mod3_stf, aes(x = Treatment, y = FoodEaten),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5,
                 position = position_dodge2(width = 1), size=1)+
  #facet_wrap(vars(Species))+
  #labs(x = "Day/Night", y = "Estimated Marginal Mean") +
  #ggtitle("Activity across the day per experiment")+
  theme_bw()

model_performance(Mass_mod3_stf)

null_mass_mod3_stf<- lmer(FoodEaten ~ 1 + (1|Ring), 
                        data= stf_Mass_data,
                na.action = na.exclude)
anova(Mass_mod3_stf, null_mass_mod3_stf, test="LTR")
```

#### FoodEaten\~Treatment\*ExperimentalDay

```{r mod4}
Mass_mod4_stf<-lmer(FoodEaten ~ Treatment*ExperimentalDay + (1|Ring),
                    data= stf_Mass_data,
                    na.action = na.exclude)

check_model(Mass_mod4_stf)

summary(Mass_mod4_stf)

mean_Mass_mod4_stf<-emmeans(Mass_mod4_stf, ~Treatment*ExperimentalDay)
mean_Mass_mod4_stf

pairs(mean_Mass_mod4_stf)

gr_Mass_mod4_stf <- ref_grid(Mass_mod4_stf, cov.keep=
                               c("Treatment", "ExperimentalDay"))
gr_emmeans_Mass_mod4_stf<-as.data.frame(emmeans(gr_Mass_mod4_stf, 
                                            specs = 
                                              c("Treatment", "ExperimentalDay"), 
                                            level=0.95))

mean_Mass_mod4_stf<- as.data.frame(mean_Mass_mod4_stf)

stf_FoodE_pval_mod2<-emmeans(gr_Mass_mod4_stf, specs = 
                              c("Treatment", "ExperimentalDay"),
                       at = list(Days = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14)),
                       level=0.95)
stf_FoodE_pval_mod2_df<-as.data.frame(pairs(stf_FoodE_pval_mod2, adjust=NULL))

raw_Mass_mod4_stf <- stf_Mass_data[c("Treatment", "FoodEaten", "ExperimentalDay")]  

plot_data_Mass_mod4_stf <- merge(gr_emmeans_Mass_mod4_stf, raw_Mass_mod4_stf,
                        by = c("Treatment", "ExperimentalDay")) 

ggplot(gr_emmeans_Mass_mod4_stf, 
       aes(x = ExperimentalDay, y = emmean, colour=Treatment)) +
  geom_jitter(data = plot_data_Mass_mod4_stf, 
              aes(x = ExperimentalDay, y = FoodEaten),
              color="lightgray") +
  geom_point() +
  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.5)+
  geom_line()+
  facet_wrap(vars(Treatment))+
  labs(x = "ExperimentalDay", y = "Estimated Marginal Mean") +
  theme_bw()

model_performance(Mass_mod4_stf)

anova(Mass_mod4_stf, Mass_mod3_stf, null_mass_mod3_stf, test="LTR")
```
